### 2018-12-09 results/analyses.RmD

---
title: "Who Needs Privacy?"
subtitle: "Analyses"
toc: TRUE
output:
  pdf_document: default
---

```{r knitr-setup, include=F}
set.seed(21)
knitr::opts_chunk$set(warning=F, echo=F, error=F, message=F,
                      cache.extra = knitr::rand_seed,
                      fig.pos = 'h')
options(digits = 3)
```

```{r r-setup, include=F}
# list packages to be used
packages <- c("corrplot", "devtools", "ggplot2", "influence.SEM", 
              "lavaan", "MBESS", "magrittr", "mice", "MVN", "PerFit", "psych", 
              "pwr", "semPlot", "tidyverse")

# install packages
install <- FALSE  # change to TRUE to install packages
if(isTRUE(install)) (
  lapply(packages, install.packages, repos = 'http://cran.us.r-project.org')
)
devtools::install_git("https://github.com/tdienlin/td")

# load packages
lapply(packages, library, character.only = TRUE)
library(td)
```

```{r data-wrangling, include=F}
# read data
# note: a bit complicated, as second row contains variable attributes
d <- read_csv("../data/data.csv", skip = 2, col_names = FALSE) %>% 
  set_names(colnames(read_csv("../data/data.csv", n_max = 1)))

# prepare table with item content
items <- data.frame(
  name = colnames(read_csv("../data/data.csv", n_max = 1)),  # extract item names
  content = colnames(read_csv("../data/data.csv", skip = 1, n_max = 2))  # extract item content
) %>% 
  mutate(content = gsub("To what extent do you agree or disagree with the following / statements[?][-]", "", content)) %>% 
  mutate(content = gsub("To what extent do you agree or disagree with the / following statements[?][-]", "", content)) %>% 
  mutate(content_breaks = sapply(content, string_break, 30))

# filter pretest data
d <- d[-(1:11), ]

# edit variables
d %<>% 
  mutate(
    male = 2 - SES.SEX,
    time = V9 - V8,
    id = 1:nrow(.)
    ) %>% 
  mutate_at(vars(RIT_1:RIT_8), funs(as.numeric(.))) %>% 
  rename(
    age = SES.AGE,
    inc = SES.JOB_2
  )

# select items
d %<>% dplyr::select(
  id, male, age, inc, time,
  starts_with("N4P"),
  starts_with("INT_"),
  SOC_1 : SOC_8,
  starts_with("FEA_"),
  TRA_1 : TRA_8,
  RIT_1 : RIT_8,
  BFI_1 : BFI_10
)

# count no of missing responses
d$miss_per <- apply(d, 1, function(x) (sum(is.na(x)))) %>% 
  divide_by(ncol(d))
```

\newpage
# Data wrangling
## Filter defective data
### Empty/missing data

```{r}
d_all <- d
n_d_all <- nrow(d_all)
table(d$miss_per) %>% round(2)

# delete respondents with more than 50% missing data
d <- d[d$miss_per < .5, ]
n_d_empty <- n_d_all - nrow(d)

# impute missing data
init <- mice(d, maxit = 0) 
vars_excl <- c("id", "time", "miss_per")
pred_ma <- init$predictorMatrix
pred_ma[, vars_excl] <- 0
d_imp_multi <- mice(d, method = "pmm", m = 1, predictorMatrix = pred_ma)
d <- mice::complete(d_imp_multi)
```

Shows that most people answered all questions. There were 15 empty data sets, and some with more than 50% missing data. These were deleted.

### Speeder

```{r speeder, warning=F}
# plot reponse times
qplot(d$time)
```

Shows that some participants took very long to answer. Inspect regular times.

```{r}
qplot(d$time[d$time < 5000])
```

Distribution looks okay (no early peak). Also no answers faster than 5 mins.

### Response patterns

```{r}
# make backup with data including suspicious patterns
d_incl_susp <- d

# run guttman tests
d$guttman <- Gpoly(dplyr::select(d, N4P.BOT_1:RIT_8) - 1, 7) %$% 
  PFscores %>% 
  unlist()

# inspect cases with susceptive data
guttman_crit <- quantile(d$guttman, probs = .95)
d_suspicious <- d[d$guttman > guttman_crit, ] %>% 
  arrange(desc(guttman))

# remove cases
d <- d[d$guttman < guttman_crit, ]
n_d_defect <- n_d_all - n_d_empty - nrow(d)
n_d <- nrow(d)
d %<>%  filter(age != 9 | is.na(age))  # filter participant with illogial age
```

All of the 5% cases indeed show extreme response patterns and/or illogical data. Will be filtered.

\newpage
# Power Analyses

```{r}
(power_analysis <- pwr.r.test(n = n_d, r = .1, sig.level = .05))
power_small <- power_analysis$power %>% multiply_by(100) %>% round(0)

(sensitivity_analysis <- pwr.r.test(n = n_d, sig.level = .05, power = .95))
sensitivity_95 <- sensitivity_analysis$r %>% round(2)
```

\newpage
# Multivariate normal distribution

```{r}
d_tmp <- dplyr::select(d, -id, -male, -age, -inc, -time, -miss_per, -guttman, -starts_with("BFI"))
(mult_norm <- mvn(data = d_tmp, mvnTest = "mardia")$multivariateNormality %>% 
    as.data.frame() %>% 
    mutate_at(vars(Statistic, "p value"), funs(as.numeric(levels(.))[.])) %>% 
    filter(Test != "MVN")
  )
```

\newpage
# Measures
## Need for Privacy

```{r, fig.height=15, fig.width=12}
d_tmp <- dplyr::select(d, starts_with("N4P"))
name_tmp <- "pri_nee"
recoded_tmp <- c("N4P.INT_1", "N4P.INT_5", "N4P.INT_8")
```

### Items

```{r, fig.height=15, fig.width=12}
ggplot(d_tmp %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# recode
d[recoded_tmp] <- 8 - d[recoded_tmp]

# new selection bc of recoding
d_tmp <- dplyr::select(d, starts_with("N4P"))

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc) 
```

\newpage
### Correlation table

```{r, fig.height=15, fig.width=12}
corr_tmp <- cor(d_tmp)
corrplot(corr_tmp, type = "upper", method = "color",
         tl.col = "black")
```

Shows that the inverted items `r recoded_tmp` do not always correlate positively and in general only weakly with the regular items. Will not be included.

\newpage
### Parallel analysis

```{r}
d_tmp <- dplyr::select(d_tmp, -one_of(recoded_tmp))
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
tmp <- omega(d_tmp, nfactors = 3, fm = "ml")
```

Shows several items with negative or low loadings, which will be excluded in updated version.

\newpage

```{r}
tmp <- omega(dplyr::select(d_tmp,
                    -N4P.BOT_4 #, -N4P.BOT_3, -N4P.INT_3, -N4P.BOT_2 #, -N4P.INT_7, -N4P.INT_4
                    ),
             nfactors = 3, fm = "ml")
```

Produces a fitting solution.

\newpage

```{r}
tmp$schmid$sl
```

Interpretation

* factor 1 measures need for privacy from the government (vertical)
* factor 2 measures need for anonymity (combined)
* factor 3 measures need for privacy from other people (horizontal)

\newpage
### CFA

```{r}
model <- '
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2 + N4P.BOT_1
  pri_nee_int =~ N4P.BOT_1 + N4P.BOT_3 + N4P.INT_4 + N4P.INT_7 + N4P.INT_9
  pri_nee_gen =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9 + N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.BOT_1 + N4P.BOT_3 + N4P.INT_4 + N4P.INT_7 + N4P.INT_9 + N4P.INT_2
  pri_nee_gen ~~ 0*pri_nee_gov + 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_gov ~~ 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_int ~~ 0*pri_nee_ano
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
```

### Table

```{r}

```

\newpage
## Integrity

```{r, fig.height=8, fig.width=12}
d_tmp <- dplyr::select(d, starts_with("INT_"))
name_tmp <- "int"
recoded_tmp <- c("INT_1", "INT_2", "INT_5", "INT_6", "INT_7", 
                 "INT_8", "INT_9", "INT_10", "INT_11")
```

### Items

```{r, fig.height=8, fig.width=12}
ggplot(d_tmp %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# recode
d[recoded_tmp] <- 8 - d[recoded_tmp]

# new selection bc of recoding
d_tmp <- dplyr::select(d, starts_with("INT_"))

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
tmp <- omega(d_tmp, fm = "ml", nfactors = 2)
```

Shows solution for two factors. Delete items INT_4 and INT_3 for not loading on general factor

\newpage

```{r}
tmp <- omega(dplyr::select(d_tmp, -INT_3, -INT_4), fm = "ml", nfactors = 2)
```

\newpage
### CFA

```{r}
model <- '
  int_1 =~ INT_1 + a*INT_5 + a*INT_6 + a*INT_2 + a*INT_7 + INT_8
  int_2 =~ b*INT_9 + b*INT_10 + b*INT_11 + b* INT_8
  int_gen =~ INT_1 + INT_2 + INT_5 + INT_6 + INT_7 + INT_8 + INT_9 + INT_10 + INT_11
  int_gen ~~ 0*int_1 + 0*int_2
  int_1 ~~ 0*int_2
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

\newpage
## Sociability
### Items

```{r, fig.height=8, fig.width=12}
d_tmp <- dplyr::select(d, SOC_1 : SOC_8)
name_tmp <- "soc"
recoded_tmp <- c("SOC_1", "SOC_3", "SOC_5", "SOC_7")

p_int <- ggplot(d_tmp %>% 
                  gather(name, value) %>% 
                  left_join(items, "name"),
                  aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)
p_int

# recode
d[recoded_tmp] <- 8 - d[recoded_tmp]

# new selection bc of recoding
d_tmp <- dplyr::select(d, SOC_1 : SOC_8)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
tmp <- omega(d_tmp, fm = "ml", nfactors = 3)
```

Shows no clear solution for three factors.

\newpage

```{r}
tmp <- omega(d_tmp, fm = "ml", nfactors = 2)
```

Shows solution for two factors. Differentiates between inverted and regular items. Delete Soc_6 and Soc_4

\newpage

```{r}
tmp <- omega(dplyr::select(d_tmp, -SOC_4, - SOC_6), fm = "ml", nfactors = 2)
```

\newpage
### CFA

```{r}
model <- '
  soc_1 =~ a*SOC_1 + SOC_3 + SOC_5 + a*SOC_7
  soc_2 =~ b*SOC_2 + b*SOC_8 + SOC_1 + SOC_7
  soc_gen =~ SOC_1 + SOC_2 + SOC_3 + SOC_5 + SOC_7 + SOC_8
  soc_gen ~~ 0*soc_1 + 0*soc_2
  soc_1 ~~ 0*soc_2
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

\newpage
## Anxiety
### Items

```{r}
d_tmp <- dplyr::select(d, starts_with("FEA_"))
name_tmp <- "fea"
recoded_tmp <- c("FEA_1", "FEA_3", "FEA_5", "FEA_7")

ggplot(d_tmp %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# recode
d[recoded_tmp] <- 8 - d[recoded_tmp]

# new selection bc of recoding
d_tmp <- dplyr::select(d, starts_with("FEA_"))

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc) 
```

### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Produces a fitting solution.

\newpage
### CFA

```{r}
model <- '
  fea_1 =~ a*FEA_2 + a*FEA_4 + a*FEA_6 + FEA_8
  fea_2 =~ FEA_1 + b*FEA_3 + b*FEA_5 + b*FEA_7
  fea_gen =~ FEA_2 + FEA_4 + FEA_7 + FEA_8 + FEA_1 + FEA_3 + FEA_5 + FEA_6

  fea_gen ~~ 0*fea_1 + 0*fea_2
  fea_1 ~~ 0*fea_2
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

\newpage
## Risk Avoidance
### Items

```{r}
d_tmp <- dplyr::select(d, RIT_1 : RIT_8)
name_tmp <- "rit"
recoded_tmp <- c("RIT_1", "RIT_3", "RIT_5")

ggplot(d_tmp %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# recode
d[recoded_tmp] <- 8 - d[recoded_tmp]

# new selection bc of recoding
d_tmp <- dplyr::select(d, RIT_1 : RIT_8)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Produces a fitting solution.

\newpage
### CFA

```{r}
model <- '
  rit_1 =~ a*RIT_2 + a*RIT_4 + a*RIT_5 + a*RIT_7 + a*RIT_8
  rit_2 =~ RIT_1 + RIT_3 + RIT_5
  rit_gen =~ RIT_1 + RIT_2 + RIT_3 + RIT_4 + RIT_5 + RIT_6 + RIT_7 + RIT_8

  rit_gen ~~ 0*rit_1 + 0*rit_2
  rit_1 ~~ 0*rit_2
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

\newpage 
## Traditionalism
### Items

```{r}
d_tmp <- dplyr::select(d, TRA_1 : TRA_8)
name_tmp <- "tra"
recoded_tmp <- c("TRA_2", "TRA_4", "TRA_6")

ggplot(d_tmp %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# recode
d[recoded_tmp] <- 8 - d[recoded_tmp]

# new selection bc of recoding
d_tmp <- dplyr::select(d, TRA_1 : TRA_8)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Need to delete item TRA_7, TRA_6, and TRA_4.

\newpage

```{r}
tmp <- omega(dplyr::select(d_tmp, -TRA_4, -TRA_7, -TRA_6, ), nfactors = 2, fm = "ml")
```

\newpage
### CFA

```{r}
model <- '
  tra_1 =~ a*TRA_1 + a*TRA_5 + a*TRA_3 + a*TRA_8 + a*TRA_7
  tra_2 =~ b*TRA_2 + b*TRA_4 + b*TRA_6
  tra_gen =~ TRA_1 + TRA_2 + TRA_3 + TRA_4 + TRA_5 + TRA_6 + TRA_7 + TRA_8

  tra_gen ~~ 0*tra_1 + 0*tra_2
  tra_1 ~~ 0*tra_2
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

\newpage

```{r}
model <- '
  tra_gen =~ TRA_1 + TRA_5 + TRA_3 + TRA_8 + TRA_7
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

\newpage
# Results

```{r}
d <- cbind(d, int_gen, soc_gen, fea_gen, tra_gen, rit_gen)
```

## Factor solution

```{r}
name_tmp <- "main"

# new
model <- '
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2 + N4P.BOT_1
  pri_nee_int =~ N4P.BOT_1 + N4P.BOT_3 + N4P.INT_4 + N4P.INT_7 + N4P.INT_9
  pri_nee_gen =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9 + N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.BOT_1 + N4P.BOT_3 + N4P.INT_4 + N4P.INT_7 + N4P.INT_9 + N4P.INT_2

  pri_nee_gen ~ int_gen + soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc
  pri_nee_gov ~ int_gen + soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc
  pri_nee_int ~ int_gen + soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc
  pri_nee_ano ~ int_gen + soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc

  int_gen ~~ soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc
  soc_gen ~~ fea_gen + tra_gen + rit_gen + male + age + inc
  fea_gen ~~ tra_gen + rit_gen + male + age + inc
  tra_gen ~~ rit_gen + male + age + inc
  rit_gen ~~ male + age + inc
  male ~~ age + inc
  age ~~ inc

  pri_nee_gen ~~ 0*pri_nee_gov + 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_gov ~~ 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_int ~~ 0*pri_nee_ano
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE, rsquare = TRUE)
assign(paste0(name_tmp, "_fit"), fit)

# fit_rem_outlier <- sem(model, filter(d, cooks_dist < 3), estimator = "MLR")
# summary(fit_rem_outlier, fit = TRUE, std = TRUE)
# d$cooks_dist <- cooks_dist
# genCookDist(model, d, estimator = "MLR")
```

\newpage
## Individual items

```{r}
name_tmp <- "main_items"
model <- '
  N4P.BOT_1 + N4P.BOT_2 + N4P.BOT_3 + N4P.BOT_4 + N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_5 + N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.SOC_9 + N4P.INT_1 + N4P.INT_2 + N4P.INT_3 + N4P.INT_4 + N4P.INT_5 + N4P.INT_6 + N4P.INT_7 + N4P.INT_8 + N4P.INT_9 ~ int_gen + soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc

  int_gen ~~ soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc
  soc_gen ~~ fea_gen + tra_gen + rit_gen + male + age + inc
  fea_gen ~~ tra_gen + rit_gen + male + age + inc
  tra_gen ~~ rit_gen + male + age + inc
  rit_gen ~~ male + age + inc
  male ~~ age + inc
  age ~~ inc
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), sem(model, d))
```

\newpage
# Tables
## Items

```{r}
t_pri_nee <- items %>% 
  filter(grepl("N4P", name)) %>% 
  mutate(name = factor(name, c(paste0("N4P.BOT_", rep(4:1)),
                             paste0("N4P.INT_", rep(9:1)),
                             paste0("N4P.SOC_", rep(9:1))))) %>% 
  arrange(desc(as.numeric(name))) %>% 
  mutate(no. = c(1:4, "--", 6:8, 5, "--", 9, 
         "--", 12, "--", "--", 13, "--", 14, 10, "--", 11, "--")) %>% 
    dplyr::select(Name = name, No. = no., Content = content)
knitr::kable(t_pri_nee) %>% 
  kableExtra::column_spec(3, width = "10cm") %>% 
  kableExtra::kable_styling(latex_options="scale_down", font_size = 7)
```

\newpage
## Psychometrics

```{r}
t_facval <- rbind(
  "Privacy need" = fit_tab(pri_nee_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Integrity" = fit_tab(int_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Sociability" = fit_tab(soc_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Anxiety" = fit_tab(fea_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Risk aversion" = fit_tab(rit_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Traditionality" = fit_tab(tra_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE)
)

t_desc <- rbind(
  pri_nee_desc, int_desc, soc_desc, fea_desc, rit_desc, tra_desc
)

(t_psyc <- cbind(t_desc, t_facval))
```

\newpage
## Results main model

```{r results}
(t_results <- coeffs_tab(main_fit, as_text = TRUE) %>% 
   mutate(Outcome = recode(Outcome, 
         `pri_nee_gen` = "Privacy need general", 
         `pri_nee_gov` = "Privacy need government",
         `pri_nee_int` = "Privacy need interpersonal",
         `pri_nee_ano` = "Privacy need anonymity")) %>% 
  mutate(Predictor = recode(Predictor,
         `int_gen` = "Integrity",
         `soc_gen` = "Sociability",
         `fea_gen` = "Anxiety",
         `rit_gen` = "Risk avoidance",
         `tra_gen` = "Traditionalism",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")))
```

\newpage
# Figures
## Bifactor model privacy needs

```{r}
p_pri_nee <- semPaths(
  pri_nee_fit,
  bifactor = "pri_nee_gen",
  layout = "tree3",
  whatLabels = "std",
  intercepts = FALSE,
  residuals = FALSE, 
  thresholds = FALSE,
  nodeLabels = c(1:14, 
           "Privacy\nneed\ngovernment", "Privacy\nneed\nanonymity", "Privacy\nneed\ninterpersonal",
           "Privacy\nneed\ngeneral"),
  reorder = FALSE,
  edge.label.position = .61,
  fixedStyle = 1,
  exoCov = FALSE,
  sizeMan = 4,
  sizeLat = 12,
  label.prop = .95,
  edge.label.cex = .8,
  mar = c(3, 1, 3, 1),
  label.scale.equal = TRUE
  )
```

\newpage
## Model bifactor

```{r}
# calculate CI for standardized effect 
# option 1: proper option, calculates extract bootstrapped coefficients for standardized solution
# CURRENTLY DOESN'T WORK
# d_effects <- bootstrapLavaan(main_fit)

# option 2: use SEs of standardized solution
d_main <- standardizedsolution(main_fit) %>% 
  as.data.frame() %>% 
  filter(op == "~") %>% 
  mutate(lhs = factor(lhs, c("pri_nee_ano", "pri_nee_int", "pri_nee_gov", "pri_nee_gen"))) %>%  # reorder for plot
  mutate(rhs = factor(rhs, c("int_gen", "soc_gen", "fea_gen", "rit_gen", "tra_gen", "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = recode(lhs, 
         `pri_nee_gen` = "Privacy need\ngeneral", 
         `pri_nee_gov` = "Privacy need\ngovernment",
         `pri_nee_int` = "Privacy need\ninterpersonal",
         `pri_nee_ano` = "Privacy need\nanonymity")) %>% 
  mutate(rhs = recode(rhs,
         `int_gen` = "Integrity",
         `soc_gen` = "Sociability",
         `fea_gen` = "Anxiety",
         `tra_gen` = "Traditio-\nnalism",
         `rit_gen` = "Risk\navoidance",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>% 
  rename(beta = est.std)

(p_main <- ggplot(d_main, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) + 
  theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = seq(-.7, .7, .2)) +
  coord_cartesian(xlim = c(-.5, .5)) +
  facet_grid(~ rhs))
```

\newpage
## Model items

```{r}
# calculate CI for standardized effect 
# option 1: proper option, calculates extract bootstrapped coefficients for standardized solution
# CURRENTLY DOESN'T WORK
# d_effects <- bootstrapLavaan(main_fit)

# option 2: use SEs of standardized solution
d_main_items <- standardizedsolution(main_items_fit) %>% 
  as.data.frame() %>% 
  filter(op == "~") %>% 
  mutate(rhs = factor(rhs, c("int_gen", "soc_gen", "fea_gen", "rit_gen", "tra_gen", "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = factor(lhs, c(paste0("N4P.BOT_", rep(4:1)),
                             paste0("N4P.INT_", rep(9:1)),
                             paste0("N4P.SOC_", rep(9:1))))) %>%  # reorder for plot
  mutate(rhs = recode(rhs,
         `int_gen` = "Integrity",
         `soc_gen` = "Sociability",
         `fea_gen` = "Anxiety",
         `tra_gen` = "Traditio-\nnalism",
         `rit_gen` = "Risk\navoidance",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>% 
  rename(beta = est.std)

(p_main_items <- ggplot(d_main_items, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) + 
  theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = seq(-.7, .7, .2)) +
  facet_grid(~ rhs))
```

\newpage
# Additional analyses
## Big five

```{r}
d_tmp <- dplyr::select(d, BFI_1 : BFI_10)
name_tmp <- "BFI"

ggplot(d_tmp %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)
```

\newpage
### Correlations

```{r}
cor(d_tmp, use = "pairwise.complete.obs")
```

The dimensions show low correlations; as a result, we will use only 1 item (the one which is not inverted), to measure each dimension. This includes the following items:

* BFI_2, BFI_3, BFI_5, BFI_8, BFI_9

```{r save-workspace}
# remove large objects
remove(d_imp_multi, d_incl_susp, fit, init)
# save workspace so that manuscript can be reproduced without rerunning analyses
save.image("../data/workspace.RData")
```