### 2019-02-12 analyses/analyses.RmD
- Major changes regarding everything: analyses (e.g., inclusion of bifactor models), ms, etc.

---
title: "Who Needs Privacy?"
subtitle: "Analyses"
output:
  pdf_document: 
    toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=F, error=F, message=F, # cache = T, 
                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
options(digits = 3, width = 100)
```

```{r r-setup, include=F}
# define packages
packages <- c("corrplot", "devtools", "GGally", "ggplot2", "kableExtra", "knitr", 
              "lavaan", "magrittr", "mice", "MVN", "PerFit", "psych", 
              "pwr", "semPlot", "semTools", "tidyverse", "td")

# install packages
# devtools::install_github("https://github.com/tdienlin/td")

# load packages
lapply(packages, library, character.only = TRUE)
```

```{r data-wrangling, include=F}
# read data
# note: a bit complicated, as second row contains variable attributes
d <- read_csv("../data/data.csv", skip = 2, col_names = FALSE) %>% 
  set_names(colnames(read_csv("../data/data.csv", n_max = 1)))

# prepare table with item content
items <- data.frame(
  name = colnames(read_csv("../data/data.csv", n_max = 1)),  # extract item names
  content = colnames(read_csv("../data/data.csv", skip = 1, n_max = 2))  # extract item content
) %>% 
  mutate(content = gsub("To what extent do you agree or disagree with the following / statements[?][-]", "", content)) %>% 
  mutate(content = gsub("To what extent do you agree or disagree with the / following statements[?][-]", "", content)) %>% 
  mutate(content_breaks = sapply(content, string_break, 30))

# edit variables
d %<>% 
  mutate_at(vars(RIT_1:RIT_8), funs(as.numeric(.))) %>% 
  mutate_at(vars(V8, V9), funs(parse_datetime(., c("%d.%m.%Y %H:%M")))) %>% 
  mutate(
    male = 2 - SES.SEX,
    time = V9 - V8,
    id = 1:nrow(.)
    ) %>% 
  rename(
    age = SES.AGE,
    inc = SES.JOB_2
  )

# select items
d %<>% dplyr::select(
  id, male, age, inc, time,
  starts_with("N4P"),
  starts_with("INT_"),
  SOC_1 : SOC_8,
  starts_with("FEA_"),
  TRA_1 : TRA_8,
  RIT_1 : RIT_8,
  BFI_1 : BFI_10
)

# count no of missing responses
d$miss_per <- apply(d, 1, function(x) (sum(is.na(x)))) %>% 
  divide_by(ncol(d))

# recode inverted items
# make dataframe of not inverted items (needed for plotting distributions)
d_noninvert <- d

# define inverted items
recoded <- c("N4P.INT_1", "N4P.INT_5", "N4P.INT_8",
             "INT_1", "INT_2", "INT_5", "INT_6", "INT_7", 
                 "INT_8", "INT_9", "INT_10", "INT_11",
             "SOC_1", "SOC_3", "SOC_5", "SOC_7",
             "FEA_1", "FEA_3", "FEA_5", "FEA_7",
             "RIT_1", "RIT_3", "RIT_5",
             "TRA_2", "TRA_4", "TRA_6")

# recode items
d[recoded] <- 8 - d[recoded]
```

\newpage
# Data wrangling
## Filter defective data
### Empty/missing data

```{r}
n_d_all <- nrow(d)

# show percentage of missing data
table(d$miss_per) %>% round(2)

# delete respondents with more than 50% missing data
d_missing <- d[d$miss_per > .5, ]
n_d_missing <- nrow(d_missing)
d <- d[!(d$id %in% d_missing$id), ]

# filter participant with illogial age
d %<>% filter(age != 9 | is.na(age))

# estimate overall percentage of missing data
na_perc <- mean(d$miss_per)

# impute missing data
init <- mice(d, maxit = 0) 
vars_excl <- c("id", "time", "miss_per")
pred_ma <- init$predictorMatrix
pred_ma[, vars_excl] <- 0
d_imp <- mice(d, method = "pmm", m = 1, maxit = 30, predictorMatrix = pred_ma, 
              seed = 180719, print = FALSE)
d <- mice::complete(d_imp)
```

Shows that most people answered all questions. There were 15 empty data sets, and some with more than 50% missing data. These were deleted. Overall, there was just `r na_perc` missing data.

### Speeder

```{r speeder, warning=F}
# plot reponse times
qplot(d$time)
```

Shows that some participants took very long to answer. Inspect regular times.

```{r}
qplot(d$time[d$time < 5000])
```

Distribution looks okay (no early peak). Also no answers faster than 5 mins.

### Response Styles

```{r}
# make backup with data including suspicious patterns
d_incl_susp <- d

# make matrixes of variables
d_pri_nee <- dplyr::select(d, starts_with("N4P"))
d_pri_nee_noninvert <- dplyr::select(d_noninvert, starts_with("N4P"))
d_int <- dplyr::select(d, starts_with("INT_"))
d_int_noninvert <- dplyr::select(d_noninvert, starts_with("INT_"))
d_soc <- dplyr::select(d, SOC_1 : SOC_8)
d_soc_noninvert <- dplyr::select(d_noninvert, SOC_1 : SOC_8)
d_fea <- dplyr::select(d, starts_with("FEA_"))
d_fea_noninvert <- dplyr::select(d_noninvert, starts_with("FEA_"))
d_rit <- dplyr::select(d, RIT_1 : RIT_8)
d_rit_noninvert <- dplyr::select(d_noninvert, RIT_1 : RIT_8)
d_tra <- dplyr::select(d, TRA_1 : TRA_8)
d_tra_noninvert <- dplyr::select(d_noninvert, TRA_1 : TRA_8)

# run guttman tests
d_vars <- list(d_pri_nee, d_int, d_soc, d_fea, d_rit, d_tra)
d_guttman_scores <- lapply(d_vars, function(x) Gnormed.poly(x - 1, 7) %$% 
  PFscores %>% 
  unlist()) %>% 
  as.data.frame() %>% 
  mutate(mean = apply(., 1, mean))
d$guttman <- d_guttman_scores$mean
ggplot(d_guttman_scores, aes(x = mean)) + 
  geom_histogram()

# inspect cases with susceptive data
# guttman_crit <- quantile(d$guttman, probs = .95)  # relative 5% criterion
guttman_crit <- .30  # absolute .25 criterion (cf. histogram)
d_suspicious <- d[d$guttman > guttman_crit, ] %>% 
  arrange(desc(guttman))

# remove cases
d <- d[d$guttman < guttman_crit, ]

# count cases
n_d_suspicious <- nrow(d_suspicious)
n_d_defect <- n_d_all - n_d_empty - nrow(d)
n_d <- nrow(d)

# make matrixes of variables (bc. of removed participants)
d_pri_nee <- dplyr::select(d, starts_with("N4P"))
d_pri_nee_noninvert <- dplyr::select(d_noninvert, starts_with("N4P"))
d_int <- dplyr::select(d, starts_with("INT_"))
d_int_noninvert <- dplyr::select(d_noninvert, starts_with("INT_"))
d_soc <- dplyr::select(d, SOC_1 : SOC_8)
d_soc_noninvert <- dplyr::select(d_noninvert, SOC_1 : SOC_8)
d_fea <- dplyr::select(d, starts_with("FEA_"))
d_fea_noninvert <- dplyr::select(d_noninvert, starts_with("FEA_"))
d_rit <- dplyr::select(d, RIT_1 : RIT_8)
d_rit_noninvert <- dplyr::select(d_noninvert, RIT_1 : RIT_8)
d_tra <- dplyr::select(d, TRA_1 : TRA_8)
d_tra_noninvert <- dplyr::select(d_noninvert, TRA_1 : TRA_8)
```

Only few participants seem to have particularly atypical data. Will filter respondents with m > .30. All of the 5% cases indeed show extreme response patterns and/or illogical data. Will be filtered.

\newpage
# Power Analyses

```{r}
(power_analysis <- pwr.r.test(n = n_d, r = .1, sig.level = .05))
power_small <- power_analysis$power %>% multiply_by(100) %>% round(0)

(power_analysis <- pwr.r.test(n = n_d, r = .2, sig.level = .05))
power_smallmoderate <- power_analysis$power %>% multiply_by(100) %>% round(0)

(sensitivity_analysis <- pwr.r.test(n = n_d, sig.level = .05, power = .95))
sensitivity_95 <- sensitivity_analysis$r %>% round(2)

(sample_analysis <- pwr.r.test(sig.level = .05, power = .95, r = .1))
sample_95 <-  sample_analysis$n %>% round(0)
```

\newpage
# Multivariate normal distribution

```{r}
d_tmp <- dplyr::select(d, -id, -male, -age, -inc, -time, -miss_per, -starts_with("BFI"))
(mult_norm <- mvn(data = d_tmp, mvnTest = "mardia")$multivariateNormality %>% 
    as.data.frame() %>% 
    mutate_at(vars(Statistic, "p value"), funs(as.numeric(levels(.))[.])) %>% 
    filter(Test != "MVN")
  )
```

\newpage
# Measures
## Need for Privacy

```{r, fig.height=15, fig.width=12}
d_tmp <- d_pri_nee
d_tmp_noninvert <- d_pri_nee_noninvert
name_tmp <- "pri_nee"
```

### Items

```{r, fig.height=15, fig.width=12}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc) 
```

\newpage
### Correlation table

```{r, fig.height=15, fig.width=12}
corr_tmp <- cor(d_tmp)
corrplot(corr_tmp, type = "upper", method = "color",
         tl.col = "black")
```

### Kaiser-Meyer-Oltkin criterion

The Kaiser-Meyer-Oltkin criterion measures the extent to which items are suitable for being combined as a single factor.

```{r}
(kmo <- KMO(d_tmp))
items_excl <- names(kmo$MSAi[kmo$MSAi < .65])
```

On the basis of the KMO, the following items should be exluded: `r names(kmo$MSAi[kmo$MSAi < .7])`.

\newpage
### Parallel analysis

```{r}
d_tmp <- dplyr::select(d_tmp, -items_excl)
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
fa(r = d_tmp, nfactors = 3, fm = "oblimin")
```

Three latent factors emerge:

- factor 1 measures need for privacy from the government (vertical)
- factor 2 measures need for privacy from other people (horizontal)
- factor 3 can be described as desire for anonymity (combined)

The following items overall contribute little:

- Communalities reveal that item INT_4 and INT_8 don't load sufficiently strong on latent factors. Should be excluded.
- Item INT_1 only loads negative on factor 3 -- little positive contribution

The following items show double-loadings:

- BOT_3, BOT_4, INT_2, INT_6. Will be difficult to decide whether to maintain or delete.

\newpage

```{r}
d_tmp <- dplyr::select(d_tmp, -c("N4P.INT_1", "N4P.INT_4", "N4P.INT_8")) # -N4P.INT_2 -N4P.INT_6, -N4P.INT_7, -N4P.BOT_2, -N4P.INT_4
fa(r = d_tmp, nfactors = 3, fm = "oblimin")
```

\newpage
### CFA 1

```{r}
model <- '
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_5 + N4P.SOC_9
  pri_nee_int =~ N4P.BOT_1 + N4P.BOT_2 + N4P.BOT_3 + N4P.BOT_4 + N4P.INT_2 + N4P.INT_3 + N4P.INT_6 + N4P.INT_7 + N4P.INT_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2 + N4P.INT_6
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Does not yield good results. Inspect modification indices.

```{r}
modificationindices(fit, minimum.value = 10, sort. = TRUE)
```

As expected, items BOT_2, BOT_3, & BOT_4 cause trouble. Will delete.

\newpage
### CFA 2

```{r}
model <- '
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_5 + N4P.SOC_9
  pri_nee_int =~ N4P.BOT_1 + N4P.INT_2 + N4P.INT_3 + N4P.INT_6 + N4P.INT_7 + N4P.INT_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2 + N4P.INT_6
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Shows acceptable fit. Problem is, we don't want to exclude too many items and to overfit the data. Let's inspect modification indices once more to see if there's a theoretically plausible adaption.

```{r}
modificationindices(fit, minimum.value = 10, sort. = TRUE)
```

Item INT_3 is a troublemaker. As it's an inverted item, we have a good reason to delete it. Also, item INT_6 doesn't really have anything to do with anonymity; we can delete it. Likewise, item SOC_5 loads on government, while it also measure anonymity. Maybe delete.

\newpage
### CFA 3

```{r}
model <- '
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9
  pri_nee_int =~ N4P.BOT_1 + N4P.INT_2 + N4P.INT_7 + N4P.INT_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_lat"), as.data.frame(predict(fit)))
```

Shows a satisfactory, but not ideal solution. Try bifactor next.

\newpage
### CFA bifactor

```{r}
model <- '
  pri_nee_gen =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9 + N4P.BOT_1 + N4P.INT_2 + N4P.INT_7 + N4P.INT_9 + N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9
  pri_nee_int =~ N4P.BOT_1 + N4P.INT_2 + N4P.INT_7 + N4P.INT_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_bifactor_fit"), fit)
```

Bifactor-solution fits the data best.

\newpage
### CFA privacy need government

```{r}
model <- '
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9
  '
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_gov_fit"), fit)
```

\newpage
### CFA privacy need interpersonal

```{r}
model <- '
  pri_nee_int =~ N4P.BOT_1 + N4P.INT_2 + N4P.INT_7 + N4P.INT_9
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_int_fit"), fit)
```

\newpage
### CFA privacy need anonymity

```{r}
model <- '
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_ano_fit"), fit)
```

\newpage
## Sociability

```{r}
d_tmp <- d_soc
d_tmp_noninvert <- d_soc_noninvert
name_tmp <- "soc"
```

### Items

```{r, fig.height=8, fig.width=12}
p_int <- ggplot(d_tmp_noninvert %>% 
                  gather(name, value) %>% 
                  left_join(items, "name"),
                  aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)
p_int

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp, fm = "ml")
```

\newpage
### EFA 1

```{r}
fa(d_tmp, fm = "ml", nfactors = 3)
```

\newpage
### EFA 2

```{r}
fa(d_tmp, fm = "ml", nfactors = 2)
```

\newpage
### CFA 1

```{r}
model <- '
  soc_1 =~ a*SOC_1 + a*SOC_3 + a*SOC_5 + a*SOC_7
  soc_2 =~ b*SOC_1 + b*SOC_2 + b*SOC_4 + b*SOC_6 + b*SOC_7 + b*SOC_8
  soc_gen =~ SOC_1 + SOC_2 + SOC_3 + SOC_4 + SOC_5 + SOC_6 + SOC_7 + SOC_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

Shows no solution for two factors.

\newpage
### CFA 2

```{r}
model <- '
  soc_gen =~ SOC_1 + SOC_2 + SOC_4 + SOC_6 + SOC_7 + SOC_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

Uni-dimensional solution with 6 items not feasible; need to reduce to 4.

\newpage
### CFA 3

```{r}
model <- '
  soc_gen =~ SOC_1 + SOC_2 + SOC_7 + SOC_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Shows adequate fit.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
## Anxiety

```{r}
d_tmp <- d_fea
d_tmp_noninvert <- d_fea_noninvert
name_tmp <- "fea"
```

### Items

```{r}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc) 
```

### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

Implies one dimension.

\newpage
### CFA 1

```{r}
model <- '
  fea_gen =~ FEA_2 + FEA_4 + FEA_7 + FEA_8 + FEA_1 + FEA_3 + FEA_5 + FEA_6
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

Doesn't fit. Instead, try two dimensions.

\newpage
### EFA

```{r}
fa(d_tmp, nfactors = 2, fm = "ml")
tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Seems appropriate.

\newpage
### CFA 2

```{r}
model <- '
  fea_1 =~ a*FEA_2 + a*FEA_4 + a*FEA_6 + a*FEA_8
  fea_2 =~ b*FEA_1 + b*FEA_3 + b*FEA_5 + b*FEA_7
  fea_gen =~ FEA_2 + FEA_4 + FEA_7 + FEA_8 + FEA_1 + FEA_3 + FEA_5 + FEA_6
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Shows an adequate solution.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
## Risk Avoidance

```{r}
d_tmp <- d_rit
d_tmp_noninvert <- d_rit_noninvert
name_tmp <- "rit"
```

### Items

```{r}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

\newpage
### EFA

```{r}
fa(d_tmp, nfactors = 2, fm = "ml")
tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Produces a fitting solution.

\newpage
### CFA 1

```{r}
model <- '
  rit_1 =~ a*RIT_2 + a*RIT_4 + RIT_5 + a*RIT_6 + a*RIT_7 + a*RIT_8
  rit_2 =~ RIT_1 + RIT_3 + RIT_5
  rit_gen =~ RIT_1 + RIT_2 + RIT_3 + RIT_4 + RIT_5 + RIT_6 + RIT_7 + RIT_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Does not produce a good-fitting model.

\newpage
### CFA 2

```{r}
model <- '
  rit_gen =~ RIT_2 + RIT_4 + RIT_6 + RIT_7 + RIT_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage 
## Traditionalism

```{r}
d_tmp <- d_tra
d_tmp_noninvert <- d_tra_noninvert
name_tmp <- "tra"
```

### Items

```{r}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp, fm = "ml")
```

\newpage
### EFA

```{r}
fa(d_tmp, nfactors = 2, fm = "ml")
tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Implies a single dimension, as on factor 2 there is only one significant loading.

\newpage
### CFA

```{r}
model <- '
  tra_gen =~ TRA_1 + TRA_5 + TRA_3 + TRA_8 + TRA_7
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Shows adequate fit.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
## Integrity

```{r, fig.height=8, fig.width=12}
d_tmp <- d_int
d_tmp_noninvert <- d_int_noninvert
name_tmp <- "int"
```

### Items

```{r, fig.height=8, fig.width=12}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### Parallel analysis

```{r}
fa.parallel(d_tmp, fm = "ml")
```

\newpage
### EFA

```{r}
fa(d_tmp, nfactors = 3, fm = "ml")
```

Two factors don't really convince, as factor 3 consists of one item only. Will try two-factor solution.

```{r}
fa(d_tmp, nfactors = 2, fm = "ml")
```

Implies that item 3 and item 4 should be dropped.

\newpage
### CFA

```{r}
model <- '
  int_1 =~ a*INT_1 + a*INT_2 + a*INT_5 + a*INT_6 + a*INT_7 + a*INT_8
  int_2 =~ b*INT_8 + b*INT_9 + b*INT_10 + b*INT_11
  int_gen =~ INT_1 + INT_2 + INT_5 + INT_6 + INT_7 + INT_8 + INT_9 + INT_10 + INT_11
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
```

Bifactor model fits the date well.

```{r}
# Extract error variance
# For the models testing the hypotheses, we next calculate the measures' error variance. The error variance will be included in the final model, enabling a better estimation of the effects.
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
# Results

```{r}
# combine model predicted values with regular data
d_lat <- cbind(soc_gen, int_gen, fea_gen, rit_gen, tra_gen, pri_nee_lat)
d <- cbind(d, d_lat)

# combine error variances in table
err_vars <- cbind(int_err_var, soc_err_var, fea_err_var, tra_err_var, rit_err_var)

# define and run model
name_tmp <- "results_threefactors"
model <- '
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9
  pri_nee_int =~ N4P.BOT_1 + N4P.INT_2 + N4P.INT_7 + N4P.INT_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2
  soc =~ soc_gen
  fea =~ fea_gen
  tra =~ tra_gen
  rit =~ rit_gen
  int =~ int_gen

  pri_nee_gov ~ soc + int + fea + tra + rit + male + age + inc
  pri_nee_int ~ soc + int + fea + tra + rit + male + age + inc
  pri_nee_ano ~ soc + int + fea + tra + rit + male + age + inc

  soc_gen ~~ .177 * soc_gen
  int_gen ~~ .104 * int_gen
  fea_gen ~~ .106 * fea_gen
  tra_gen ~~ .169 * tra_gen
  rit_gen ~~ .118 * rit_gen
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
```

\newpage
# Tables
## Items measuring need for privacy

```{r}
# fig.pos="h"
# get names of included items
items_incl <- colnames(inspect(results_threefactors_fit, 'sampstat')$cov)
pri_nee_gov_items <- parameterestimates(results_threefactors_fit) %>% 
  filter(lhs == "pri_nee_gov", op == "=~") %>% 
  select(rhs) %>% 
  unlist()
pri_nee_int_items <- parameterestimates(results_threefactors_fit) %>% 
  filter(lhs == "pri_nee_int", op == "=~") %>% 
  select(rhs) %>% 
  unlist()
pri_nee_ano_items <- parameterestimates(results_threefactors_fit) %>% 
  filter(lhs == "pri_nee_ano", op == "=~") %>% 
  select(rhs) %>% 
  unlist()

t_pri_nee <- items %>% 
  filter(grepl("N4P", name)) %>% 
  mutate(name = factor(name, c(paste0("N4P.BOT_", rep(4:1)),
                             paste0("N4P.INT_", rep(9:1)),
                             paste0("N4P.SOC_", rep(9:1))))) %>% 
  mutate(included = ifelse(name %in% items_incl, "yes", "no")) %>% 
  arrange(desc(as.numeric(name))) %>% 
  dplyr::select(name, included, content)

knitr::kable(t_pri_nee) %>% 
  kableExtra::column_spec(3, width = "10cm") %>% 
  kableExtra::kable_styling(latex_options="scale_down", font_size = 7)
```

\newpage
## Psychometrics

```{r}
t_facval <- rbind(
  "(Combined)" = fit_tab(pri_nee_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Government" = fit_tab(pri_nee_gov_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Interpersonal" = fit_tab(pri_nee_int_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Anonymity" = fit_tab(pri_nee_ano_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Sociability" = fit_tab(soc_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Integrity" = fit_tab(int_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Anxiety" = fit_tab(fea_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Risk aversion" = fit_tab(rit_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Traditionality" = fit_tab(tra_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE)
)

pri_nee_gov_desc <- c(mean = mean(as.matrix(d[pri_nee_gov_items])),
                      sd = sd(as.matrix(d[pri_nee_gov_items])))
pri_nee_int_desc <- c(mean = mean(as.matrix(d[pri_nee_int_items])),
                      sd = sd(as.matrix(d[pri_nee_int_items])))
pri_nee_ano_desc <- c(mean = mean(as.matrix(d[pri_nee_ano_items])),
                      sd = sd(as.matrix(d[pri_nee_ano_items])))

t_desc <- rbind(pri_nee_desc, pri_nee_gov_desc, pri_nee_int_desc, pri_nee_ano_desc,
                soc_desc, int_desc, fea_desc, rit_desc, tra_desc)

t_psyc <- cbind(t_desc, t_facval)
kable(t_psyc)
```

\newpage
## Results

```{r}
t_results_threefactors <- coeffs_tab(results_threefactors_fit, as_text = TRUE) %>% 
   mutate(Outcome = recode(Outcome, 
         `pri_nee_gov` = "Privacy need government",
         `pri_nee_int` = "Privacy need interpersonal",
         `pri_nee_ano` = "Privacy need anonymity")) %>% 
  mutate(Predictor = recode(Predictor,
         `int` = "Integrity",
         `soc` = "Sociability",
         `fea` = "Anxiety",
         `rit` = "Risk avoidance",
         `tra` = "Traditionalism",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male"))
kable(t_results_threefactors)
```

\newpage
# Figures
## Bivariate relations

```{r eval = F}
d_lat %<>% set_names("Sociability", "Integrity", "Anxiety", "Risk\navoidance", "Traditio-\nnalism", "Privacy need\ngovernment", "Privacy need\ninterpersonal", "Privacy need\nanonymity")

p_bivar <- ggpairs(
  d_lat,
  upper = list(continuous = "cor"),
  lower = list(continuous = scat_plot)) +
  theme_bw()
print(p_bivar)
```

## Results

```{r}
# calculate CI for standardized effect 
d_results_threefactors <- standardizedsolution(results_threefactors_fit) %>% 
  as.data.frame() %>% 
  filter(op == "~") %>% 
  mutate(lhs = factor(lhs, c("pri_nee_ano", "pri_nee_int", "pri_nee_gov"))) %>%  # reorder for plot
  mutate(rhs = factor(rhs, c("int", "soc", "fea", "rit", "tra", "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = recode(lhs, 
         `pri_nee_gov` = "Privacy need\ngovernment",
         `pri_nee_int` = "Privacy need\ninterpersonal",
         `pri_nee_ano` = "Privacy need\nanonymity")) %>% 
  mutate(rhs = recode(rhs,
         `int` = "Integrity",
         `soc` = "Sociability",
         `fea` = "Anxiety",
         `tra` = "Traditio-\nnalism",
         `rit` = "Risk\navoidance",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>% 
  rename(beta = est.std)

(p_results_threefactors <- ggplot(d_results_threefactors, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) + 
  theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = c(-.4, -.2, 0, .2, .4)) +
  coord_cartesian(xlim = c(-.5, .5)) +
  facet_grid(~ rhs))
```

\newpage
# Additional analyses
## Bi-Factor

```{r}
name_tmp <- "results_bifactor"
model <- '
  pri_nee_gen =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9 + N4P.BOT_1 + N4P.INT_2 + N4P.INT_7 + N4P.INT_9 + N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2
  pri_nee_gov =~ N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_9
  pri_nee_int =~ N4P.BOT_1 + N4P.INT_2 + N4P.INT_7 + N4P.INT_9
  pri_nee_ano =~ N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.INT_2

  int =~ int_gen
  soc =~ soc_gen
  fea =~ fea_gen
  tra =~ tra_gen
  rit =~ rit_gen

  pri_nee_gen ~ int + soc + fea + tra + rit + male + age + inc
  pri_nee_gov ~ int + soc + fea + tra + rit + male + age + inc
  pri_nee_int ~ int + soc + fea + tra + rit + male + age + inc
  pri_nee_ano ~ int + soc + fea + tra + rit + male + age + inc

  pri_nee_gen ~~ 0*pri_nee_gov + 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_gov ~~ 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_int ~~ 0*pri_nee_ano

  int_gen ~~ .104 * int_gen
  soc_gen ~~ .177 * soc_gen
  fea_gen ~~ .106 * fea_gen
  tra_gen ~~ .169 * tra_gen
  rit_gen ~~ .118 * rit_gen
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE, rsquare = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
```

\newpage

```{r}
# calculate CI for standardized effect
d_results_bifactor <- standardizedsolution(results_bifactor_fit) %>%
  as.data.frame() %>%
  filter(op == "~") %>%
  mutate(lhs = factor(lhs, c("pri_nee_ano", "pri_nee_int", "pri_nee_gov", "pri_nee_gen"))) %>%  # reorder for plot
  mutate(rhs = factor(rhs, c("int", "soc", "fea", "rit", "tra", "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = recode(lhs,
         `pri_nee_gen` = "Privacy need\ngeneral",
         `pri_nee_gov` = "Privacy need\ngovernment",
         `pri_nee_int` = "Privacy need\ninterpersonal",
         `pri_nee_ano` = "Privacy need\nanonymity")) %>%
  mutate(rhs = recode(rhs,
         `int` = "Integrity",
         `soc` = "Sociability",
         `fea` = "Anxiety",
         `tra` = "Traditio-\nnalism",
         `rit` = "Risk\navoidance",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>%
  rename(beta = est.std)

(p_results_bifactor <- ggplot(d_results_bifactor, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) +
  theme_bw() +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = c(-.4, -.2, 0, .2, .4)) +
  coord_cartesian(xlim = c(-.5, .5)) +
  facet_grid(~ rhs))
```

\newpage
## Individual items

```{r}
name_tmp <- "results_items"
model <- '
  N4P.BOT_1 + N4P.BOT_2 + N4P.BOT_3 + N4P.BOT_4 + N4P.SOC_1 + N4P.SOC_2 + N4P.SOC_3 + N4P.SOC_4 + N4P.SOC_5 + N4P.SOC_6 + N4P.SOC_7 + N4P.SOC_8 + N4P.SOC_9 + N4P.INT_1 + N4P.INT_2 + N4P.INT_3 + N4P.INT_4 + N4P.INT_5 + N4P.INT_6 + N4P.INT_7 + N4P.INT_8 + N4P.INT_9 ~ int_gen + soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc

  int_gen ~~ soc_gen + fea_gen + tra_gen + rit_gen + male + age + inc
  soc_gen ~~ fea_gen + tra_gen + rit_gen + male + age + inc
  fea_gen ~~ tra_gen + rit_gen + male + age + inc
  tra_gen ~~ rit_gen + male + age + inc
  rit_gen ~~ male + age + inc
  male ~~ age + inc
  age ~~ inc
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), sem(model, d))
```

\newpage

```{r}
# calculate CI for standardized effect 
d_results_items <- standardizedsolution(results_items_fit) %>% 
  as.data.frame() %>% 
  filter(op == "~") %>% 
  mutate(rhs = factor(rhs, c("int_gen", "soc_gen", "fea_gen", "rit_gen", "tra_gen", 
                             "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = factor(lhs, c(paste0("N4P.BOT_", rep(4:1)),
                             paste0("N4P.INT_", rep(9:1)),
                             paste0("N4P.SOC_", rep(9:1))))) %>%  # reorder for plot
  mutate(rhs = recode(rhs,
         `int_gen` = "Integrity",
         `soc_gen` = "Sociability",
         `fea_gen` = "Anxiety",
         `tra_gen` = "Traditio-\nnalism",
         `rit_gen` = "Risk\navoidance",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>% 
  rename(beta = est.std)

(p_results_items <- ggplot(d_results_items, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) + 
  theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = seq(-.7, .7, .2)) +
  facet_grid(~ rhs))
```

```{r save-workspace}
# remove large objects
remove(d_imp, d_incl_susp, fit)

# save workspace so that manuscript can be reproduced without rerunning analyses
save.image("../data/workspace.RData")
```