### 2019-02-26 analyses/analyses.RmD
Major updates on everything. Specifically, this includes updates on data, including MMs latest feedback, and getting things ready for submission.

---
title: "Who Needs Privacy?"
subtitle: "Analyses"
output:
  pdf_document: 
    toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=F, error=F, message=F, # cache = T, 
                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
options(digits = 3, width = 90)
```

```{r r-setup, include=F}
# define packages
packages <- c("corrplot", "devtools", "GGally", "ggplot2", "kableExtra", "knitr", 
              "lavaan", "magrittr", "mice", "MVN", "PerFit", "psych", 
              "pwr", "semPlot", "semTools", "tidyverse", "td")

# install packages
devtools::install_github("https://github.com/tdienlin/td")

# load packages
lapply(packages, library, character.only = TRUE)
```

\newpage
# Items

"To what extent do you agree or disagree with the following statements?"

(-3) Strongly Disagree, (-2) Disagree, (-1) Slightly Disagree, (0) Neutral, (1) Slightly Agree (2) Agree, (3) Strongly Agree

(Inverted items are labelled "*".)

## Need for privacy
### Informational

1.	I prefer it when other people do not know much about me.
2.	When given the chance, I prefer being incognito.
3.	I don't want personal information about me being publicly available.
4.	Not everybody needs to know everything about me.

### Societal

Government Surveillance

1.	I need government agencies to respect my privacy, even if that hinders a greater societal cause.
2.	I need the information that companies (e.g., Amazon, Facebook, or Google) have about me to stay private so that the government cannot access it.
3.	I don't want the government to gather information about me, even if that makes it more difficult for them to spend tax income efficiently.
4.	I don't want government agencies to monitor my personal communication, even if doing so prevents future terrorist attacks.

Anonymity

5.	I need to be able to surf online anonymously.
6.	I need to be able to use a fake name on social network sites to preserve my privacy.
7.	I feel the need to avoid places with video surveillance.
8.	I prefer not to carry my ID with me all the time to preserve my privacy

Both

9.	I feel the need to protect my privacy from government agencies.

### Interpersonal

Online

1.	I feel the need to disclose personal information about me on social network sites.
2.	My need for privacy is so strong that it prevents me from using Facebook actively.
3.	I don’t feel the need to be able to communicate about very personal things with others online.
4.	I need to know that my boss or future employers cannot find information about me online that they might disapprove of.

Offline

5.	I always need a person to talk about personal things.*
6.	I don’t need to know a lot of things about people I interact with, as that might cause problems.
7.	I don’t feel the need to tell my friends all my secrets.
8.	I sometimes feel the need to share my personal point of view with someone I don’t know that well.*

Both

9.	I feel the need to protect my privacy from other people.

## Sociability

1.	I shy away from crowds of people.*
2.	I like to have a lot of people around me.
3.	I usually prefer to do things alone.*
4.	I really feel the need for other people if I am by myself for long.
5.	I prefer jobs that let me work alone without being bothered by other people.*
6.	I’d rather vacation at a popular beach than an isolated cabin in the woods.
7.	Social gatherings are usually boring to me.*
8.	I enjoy parties with lots of people.

## Integrity

1.	I am often tempted to take things that do not belong to me.*
2.	I don't think there's anything wrong with cheating a little on one's income tax forms.*
3.	When traveling, I would always declare everything at customs.
4.	If I found a wallet, I would not be tempted to keep any money.
5.	If I could get away with it, I would leave a restaurant or bar without paying the bill.*
6.	I would buy stolen merchandise if the price was right.*
7.	I would mislead people when it comes to my academic achievements.*
8.	I think that there are circumstances when people are justified in cheating.*

Self-designed

9.	I have cheated on an exam in school.*
10.	I have watched movies or listened to music online for free, even though I knew doing so violated copyright law.*
11.	Certain things that I have done would be considered illegal in some places.*

## Anxiety

1.	I am not a worrier.*
2.	I am easily frightened.
3.	I rarely feel fearful or anxious.*
4.	I often feel tense and jittery.
5.	I’m seldom apprehensive about the future.
6.	I often worry about things that might go wrong.
7.	I have fewer fears than most people.*
8.	Frightening thoughts sometimes come into my head.

## Risk avoidance

1.	Over the years I’ve done some pretty stupid things.*
2.	I think things through before coming to a decision.
3.	Occasionally I act first and think later.*
4.	I always consider the consequences before I take action.
5.	I often do things on the spur of the moment.*
6.	I rarely make hasty decisions.
7.	I plan ahead carefully when I go on a trip.
8.	I think twice before I answer a question.

## Traditionalism

1.	I'm pretty set in my ways.
2.	I think it's interesting to learn and develop new hobbies.*
3.	Once I find the right way to do something, I stick to it.
4.	I often try new and foreign foods.*
5.	I prefer to spend my time in familiar surroundings.
6.	Sometimes I make changes around the house just to try something different.*
7.	On a vacation, I prefer going back to a tried and true spot.
8.	I follow the same route when I go someplace.

\newpage
# Sample

```{r data-wrangling, include=F}
# read data
# note: a bit complicated, as second row contains variable attributes
d <- read_csv("../data/data.csv", skip = 2, col_names = FALSE) %>% 
  set_names(colnames(read_csv("../data/data.csv", n_max = 1)))
items <- read_csv("../data/items.csv")

# count no of missing responses
d$miss_per <- apply(d, 1, function(x) (sum(is.na(x)))) %>% 
  divide_by(ncol(d))

# recode inverted items
# make dataframe of not inverted items (needed for plotting distributions)
d_noninvert <- d

# define inverted items
recoded <- c("pri_nee_int_1", "pri_nee_int_5", "pri_nee_int_8",
             "itg_1", "itg_2", "itg_5", "itg_6", "itg_7", 
                 "itg_8", "itg_9", "itg_10", "itg_11",
             "soc_1", "soc_3", "soc_5", "soc_7",
             "anx_1", "anx_3", "anx_5", "anx_7",
             "ria_1", "ria_3", "ria_5",
             "tra_2", "tra_4", "tra_6")

# recode items
d[recoded] <- 8 - d[recoded]
```

## Missing values

Inspect how much data is missing.

```{r}
n_d_all <- nrow(d)

# show percentage of missing data
table(d$miss_per) %>% round(2)

# delete respondents with more than 50% missing data
d_missing <- d[d$miss_per > .5, ]
n_d_missing <- nrow(d_missing)
d <- d[!(d$id %in% d_missing$id), ]

# filter participant with illogial age
d %<>% filter(age != 9 | is.na(age))

# estimate overall percentage of missing data
na_perc <- mean(d$miss_per)

# impute missing data
init <- mice(d, maxit = 0) 
vars_excl <- c("id", "time", "miss_per")
pred_ma <- init$predictorMatrix
pred_ma[, vars_excl] <- 0
d_imp <- mice(d, method = "pmm", m = 1, maxit = 30, predictorMatrix = pred_ma, 
              seed = 180719, print = FALSE)
d <- mice::complete(d_imp)
```

Shows that most people answered all questions. There were 15 empty data sets, and some with more than 50% missing data. These were deleted. In the final sample, there was `r na_perc` missing data.

## Response time

Inspect how long it took respondents to answer the questionnaire.

```{r speeder, warning=F}
# plot reponse times
qplot(d$time)
```

Shows that some participants took very long to answer. Inspect regular times.

```{r}
qplot(d$time[d$time < 5000])
```

Distribution looks okay (no early peak). Also no answers faster than 5 mins. No respondents will be excluded.

\newpage
## Response patterns

We inspect response patterns using the Guttman criterion. Cases with most extreme values will be excluded.

```{r}
# make backup with data including suspicious patterns
d_incl_susp <- d

# make matrixes of variables
d_pri_nee <- dplyr::select(d, starts_with("pri_nee"))
d_pri_nee_noninvert <- dplyr::select(d_noninvert, starts_with("pri_nee"))
d_itg <- dplyr::select(d, starts_with("itg_"))
d_itg_noninvert <- dplyr::select(d_noninvert, starts_with("itg_"))
d_soc <- dplyr::select(d, soc_1 : soc_8)
d_soc_noninvert <- dplyr::select(d_noninvert, soc_1 : soc_8)
d_anx <- dplyr::select(d, starts_with("anx_"))
d_anx_noninvert <- dplyr::select(d_noninvert, starts_with("anx_"))
d_ria <- dplyr::select(d, ria_1 : ria_8)
d_ria_noninvert <- dplyr::select(d_noninvert, ria_1 : ria_8)
d_tra <- dplyr::select(d, tra_1 : tra_8)
d_tra_noninvert <- dplyr::select(d_noninvert, tra_1 : tra_8)

# run guttman tests
d_vars <- list(d_pri_nee, d_itg, d_soc, d_anx, d_ria, d_tra)
d_guttman_scores <- lapply(d_vars, function(x) Gnormed.poly(x - 1, 7) %$% 
  PFscores %>% 
  unlist()) %>% 
  as.data.frame() %>% 
  mutate(mean = apply(., 1, mean))
d$guttman <- d_guttman_scores$mean
ggplot(d_guttman_scores, aes(x = mean)) + 
  geom_histogram()

# inspect cases with susceptive data
guttman_crit_30 <- .30  # absolute .30 criterion (cf. histogram)
guttman_crit_95 <- quantile(d$guttman, .95)  # relative 5% criterion; possible, but given the already small sample not recommended

d_suspicious <- d[d$guttman > guttman_crit_30, ] %>% 
  arrange(desc(guttman))
n_d_suspicious <- nrow(d_suspicious)

# remove cases
d <- d[d$guttman < guttman_crit_30, ]
d_guttman_95 <- d[d$guttman < guttman_crit_95, ]

# count cases
n_d <- nrow(d)

# make matrixes of variables (bc. of removed participants)
d_pri_nee <- dplyr::select(d, starts_with("pri_nee"))
d_pri_nee_noninvert <- dplyr::select(d_noninvert, starts_with("pri_nee"))
d_itg <- dplyr::select(d, starts_with("itg_"))
d_itg_noninvert <- dplyr::select(d_noninvert, starts_with("itg_"))
d_soc <- dplyr::select(d, soc_1 : soc_8)
d_soc_noninvert <- dplyr::select(d_noninvert, soc_1 : soc_8)
d_anx <- dplyr::select(d, starts_with("anx_"))
d_anx_noninvert <- dplyr::select(d_noninvert, starts_with("anx_"))
d_ria <- dplyr::select(d, ria_1 : ria_8)
d_ria_noninvert <- dplyr::select(d_noninvert, ria_1 : ria_8)
d_tra <- dplyr::select(d, tra_1 : tra_8)
d_tra_noninvert <- dplyr::select(d_noninvert, tra_1 : tra_8)
```

Only few participants seem to have particularly atypical data. Will filter respondents with m > .30. Inspect those cases to see whether they indeed show irregular response patterns.

\newpage

```{r}
d_suspicious
```

The cases indeed show extreme response patterns. Also, several inverted variables were not recognized (e.g., `r recoded`). These respondents will be deleted.

\newpage
# Power analyses

Estimate achieved power for small effects.

```{r}
(power_analysis <- pwr.r.test(n = n_d, r = .1, sig.level = .05))
power_small <- power_analysis$power %>% multiply_by(100) %>% round(0)
```

Estimate achieved power for small to moderate effects.

```{r}
(power_analysis <- pwr.r.test(n = n_d, r = .2, sig.level = .05))
power_smallmoderate <- power_analysis$power %>% multiply_by(100) %>% round(0)
```

Estimate the size of effects that we were able to find with a probability of 95%.

```{r}
(sensitivity_analysis <- pwr.r.test(n = n_d, sig.level = .05, power = .95))
sensitivity_95 <- sensitivity_analysis$r %>% round(2)
```

Estimate the sample size needed to find small effects with a probability of 95%.

```{r}
(sample_analysis <- pwr.r.test(sig.level = .05, power = .95, r = .1))
sample_95 <-  sample_analysis$n %>% round(0)
```

\newpage
# Multivariate normal distribution

We test for the assumption of multivariate normal distribution by running Mardia's test.

```{r}
d_tmp <- dplyr::select(d, -id, -male, -age, -inc, -time, -miss_per)
(mult_norm <- mvn(data = d_tmp, mvnTest = "mardia")$multivariateNormality %>% 
    as.data.frame() %>% 
    mutate_at(vars(Statistic, "p value"), funs(as.numeric(levels(.))[.])) %>% 
    filter(Test != "MVN")
  )
```

Shows that the assumptions of multivariate normal distribution was violated. Will hence use robust estimator in the following analyses.

\newpage
# Measures

In what follows, we present the measures we used. Specifically, we present each item's distribution and a confirmatory factor analysis (CFA) of the original scale. If the scale did not show sufficient fit, we first ran exploratory factor analyses (EFA) to better determine the underlying factor structure. In an iterative process, we then adapted the scales (e.g., by introducing subdimensions, trying bifactor solutions, or deleting items).

## Need for Privacy

```{r}
d_tmp <- d_pri_nee
d_tmp_noninvert <- d_pri_nee_noninvert
name_tmp <- "pri_nee"
```

### Items

```{r, fig.height=15, fig.width=12}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc) 
```

\newpage
### CFA 1

We first analyze all dimensions separately to see how they do individually and as specified theoretically a priori. The general aim is, of course, to find the general structure of the items when analyzed together.

#### Informational Privacy

```{r}
model <- '
  pri_nee_gen =~ pri_nee_gen_1 + pri_nee_gen_2 + pri_nee_gen_3 + pri_nee_gen_4
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Does not show adequate fit.

\newpage
#### Government

```{r}
model <- '
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_5 + pri_nee_soc_9
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Shows good fit, can be used individually for future research.

\newpage
#### Anonymity

```{r}
model <- '
  pri_nee_ano =~ pri_nee_soc_5 + pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Shows moderate fit.

\newpage
#### Interpersonal online

```{r}
model <- '
  pri_nee_int_onl =~ pri_nee_int_1 + pri_nee_int_2 + pri_nee_int_3 + pri_nee_int_4 + pri_nee_int_9
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Shows poor fit.

\newpage
#### Interpersonal offline

```{r}
model <- '
  pri_nee_int_off =~ pri_nee_int_5 + pri_nee_int_6 + pri_nee_int_7 + pri_nee_int_8 + pri_nee_int_9
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Does not show adequate fit.

\newpage
#### Combined

```{r}
model <- '
  pri_nee_gen =~ pri_nee_gen_1 + pri_nee_gen_2 + pri_nee_gen_3 + pri_nee_gen_4
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_5 + pri_nee_soc_9
  pri_nee_ano =~ pri_nee_soc_5 + pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8
  pri_nee_int_onl =~ pri_nee_int_1 + pri_nee_int_2 + pri_nee_int_3 + pri_nee_int_4 + pri_nee_int_9
  pri_nee_int_off =~ pri_nee_int_5 + pri_nee_int_6 + pri_nee_int_7 + pri_nee_int_8 + pri_nee_int_9
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

The combined scale does not converge. Will need to be adapted thoroughly.

\newpage
### EFA 1

In what follows, we now try to find what items need to be selected in order to attain a measure of need for privacy that includes as many items as possible.

#### Kaiser-Meyer-Oltkin criterion

The Kaiser-Meyer-Oltkin criterion measures the extent to which items are suitable for being combined as a single factor.

```{r}
(kmo <- KMO(d_tmp))
items_excl <- names(kmo$MSAi[kmo$MSAi < .65])
```

On the basis of the KMO, the following items should (and will) be exluded: `r names(kmo$MSAi[kmo$MSAi < .7])`.

\newpage
#### Parallel analysis

Next, we run a parallel analysis to determine the underlying structure.

```{r}
d_tmp <- dplyr::select(d_tmp, -items_excl)
fa.parallel(d_tmp)
```

\newpage
#### Factor analysis 1

```{r}
efa <- fa(r = d_tmp, nfactors = 3, fm = "oblimin")
efa$loadings
```

Three latent factors emerge:

- factor 1 measures need for privacy from the government (vertical)
- factor 2 measures need for privacy from other people (horizontal)
- factor 3 can be described as desire for anonymity (combined)

The following items will be exluded:

- Communalities reveal that item itg_4 and itg_8 don't load sufficiently strong on latent factors.
- Item itg_1 only loads negative on factor 3 -- little positive contribution.

The following items show double-loadings:

- BOT_3, BOT_4, itg_2, itg_6. Will be difficult to decide whether to maintain or delete.

Run novel factor analysis.

\newpage
#### Factor analysis 2

```{r}
d_tmp <- dplyr::select(d_tmp, -c("pri_nee_int_1", "pri_nee_int_4", "pri_nee_int_8"))
efa <- fa(r = d_tmp, nfactors = 3, fm = "oblimin")
efa$loadings
```

Looks better. Will now be tested in CFA.

\newpage
### CFA 2

```{r}
model <- '
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_5 + pri_nee_soc_9
  pri_nee_int =~ pri_nee_gen_1 + pri_nee_gen_2 + pri_nee_gen_3 + pri_nee_gen_4 + pri_nee_int_2 + pri_nee_int_3 + pri_nee_int_6 + pri_nee_int_7 + pri_nee_int_9
  pri_nee_ano =~ pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2 + pri_nee_int_6
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Does not yield good results. Inspect modification indices.

```{r}
modificationindices(fit, minimum.value = 10, sort. = TRUE)
```

As expected, items BOT_2, BOT_3, & BOT_4 cause trouble. Will be deleted.

\newpage
### CFA 3

```{r}
model <- '
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_5 + pri_nee_soc_9
  pri_nee_int =~ pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_3 + pri_nee_int_6 + pri_nee_int_7 + pri_nee_int_9
  pri_nee_ano =~ pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2 + pri_nee_int_6
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Shows shows improved but still not really acceptable fit. Problem is, we don't want to exclude too many items and to overfit the data. Let's inspect modification indices once more to see if there's a theoretically plausible adaption.

```{r}
modificationindices(fit, minimum.value = 10, sort. = TRUE)
```

Item itg_3 is a troublemaker. As it's an inverted item, we have a good reason to delete it. Also, item itg_6 doesn't really have anything to do with anonymity; we can delete it. Likewise, item soc_5 loads on government, while it also measure anonymity. Maybe delete.

\newpage
### CFA 4

```{r}
model <- '
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_9
  pri_nee_int =~ pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_7 + pri_nee_int_9
  pri_nee_ano =~ pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_lat"), as.data.frame(predict(fit)))
```

Although not ideal, shows a satisfactory solution. Will be used for analyses. In what follows, run specific CFAs for the dimensions, to get a better idea how they would work when used separately.

\newpage
#### CFA privacy need government

```{r}
model <- '
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_9
  '
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_gov_fit"), fit)
```

\newpage
#### CFA privacy need interpersonal

```{r}
model <- '
  pri_nee_int =~ pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_7 + pri_nee_int_9
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_int_fit"), fit)
```

\newpage
#### CFA privacy need anonymity

```{r}
model <- '
  pri_nee_ano =~ pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_ano_fit"), fit)
```

\newpage
#### CFA bifactor

```{r}
model <- '
  pri_nee_gen =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_9 + pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_7 + pri_nee_int_9 + pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_9
  pri_nee_int =~ pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_7 + pri_nee_int_9
  pri_nee_ano =~ pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_bifactor_fit"), fit)
```

Bifactor-solution fits the data best. However, for theoretical reasons will not be used for the analyses.

\newpage
## Sociability

```{r}
d_tmp <- d_soc
d_tmp_noninvert <- d_soc_noninvert
name_tmp <- "soc"
```

### Items

```{r, fig.height=8, fig.width=12}
ggplot(d_tmp_noninvert %>% 
                  gather(name, value) %>% 
                  left_join(items, "name"),
                  aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### CFA 1

```{r}
model <- '
  soc =~ soc_1 + soc_2 + soc_3 + soc_4 + soc_5 + soc_6 + soc_7 + soc_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

Does not show sufficient fit. Run EFAs to determine underlying factor structure.

\newpage
### EFA 1
#### Parallel analysis

```{r}
fa.parallel(d_tmp, fm = "ml")
```

\newpage
#### Factor analysis

```{r}
efa <- fa(d_tmp, fm = "ml", nfactors = 3)
efa$loadings
```

The three factor solution does not show a coherent picture; for example, Factor 3 is determined by a single item.

\newpage
### EFA 2
#### Factor analysis

```{r}
efa <- fa(d_tmp, fm = "ml", nfactors = 2)
efa$loadings
```

Seems more plausible. Will be modelled as bifactor solution next.

\newpage
### CFA 2

```{r}
model <- '
  soc_one =~ a*soc_1 + a*soc_3 + a*soc_5 + a*soc_7
  soc_two =~ b*soc_1 + b*soc_2 + b*soc_4 + b*soc_6 + b*soc_7 + b*soc_8
  soc_gen =~ soc_1 + soc_2 + soc_3 + soc_4 + soc_5 + soc_6 + soc_7 + soc_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

Shows no solution for two factors. Single-dimension solution will be fitted next.

\newpage
### CFA 3

```{r}
model <- '
  soc_gen =~ soc_1 + soc_2 + soc_4 + soc_6 + soc_7 + soc_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

Uni-dimensional solution with 6 items not feasible; need to reduce to 4.

\newpage
### CFA 4

```{r}
model <- '
  soc_gen =~ soc_1 + soc_2 + soc_7 + soc_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Shows adequate fit.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
## Integrity

```{r, fig.height=8, fig.width=12}
d_tmp <- d_itg
d_tmp_noninvert <- d_itg_noninvert
name_tmp <- "itg"
```

### Items

```{r, fig.height=8, fig.width=12}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

Does not show adequate fit.

\newpage
### CFA 1

```{r}
model <- '
  itg =~ itg_1 + itg_2 + itg_3 + itg_4 + itg_5 + itg_6 + itg_7 + itg_8 + itg_9 + itg_10 + itg_11
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

\newpage
### EFA 1
#### Parallel analysis

```{r}
fa.parallel(d_tmp, fm = "ml")
```

\newpage
#### Factor analysis 1

```{r}
efa <- fa(d_tmp, nfactors = 3, fm = "ml")
efa$loadings
```

Two factors don't really convince, as factor 3 consists of one item only. Will try two-factor solution.

\newpage
#### Factor analysis 2

```{r}
efa <- fa(d_tmp, nfactors = 2, fm = "ml")
efa$loadings
```

Suggests that item 3 and item 4 should be dropped.

\newpage
### CFA 2

```{r}
model <- '
  int_1 =~ a*itg_1 + a*itg_2 + a*itg_5 + a*itg_6 + a*itg_7 + a*itg_8
  int_2 =~ b*itg_8 + b*itg_9 + b*itg_10 + b*itg_11
  itg_gen =~ itg_1 + itg_2 + itg_5 + itg_6 + itg_7 + itg_8 + itg_9 + itg_10 + itg_11
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
```

The data fit a bifactor model well.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
## Anxiety

```{r}
d_tmp <- d_anx
d_tmp_noninvert <- d_anx_noninvert
name_tmp <- "anx"
```

### Items

```{r}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc) 
```

\newpage
### CFA 1

```{r}
model <- '
  anx =~ anx_1 + anx_2 + anx_3 + anx_4 + anx_5 + anx_6 + anx_7 + anx_8
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
```

Does not show adequate fit.

\newpage
### EFA 1
#### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

Implies three dimension.

\newpage
#### Factor analysis

```{r}
efa <- fa(d_tmp, nfactors = 2, fm = "ml")
efa$loadings
```

Seems appropriate. First, try bifactor solution.

\newpage
### CFA 2

```{r}
model <- '
  anx_one =~ a*anx_2 + a*anx_4 + a*anx_6 + a*anx_8
  anx_two =~ b*anx_1 + b*anx_3 + b*anx_5 + b*anx_7
  anx_gen =~ anx_2 + anx_4 + anx_7 + anx_8 + anx_1 + anx_3 + anx_5 + anx_6
'
fit <- cfa(model, d, estimator = "MLM", orthogonal = TRUE)
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Shows an adequate solution.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
## Risk avoidance

```{r}
d_tmp <- d_ria
d_tmp_noninvert <- d_ria_noninvert
name_tmp <- "ria"
```

### Items

```{r}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### CFA 1

```{r}
model <- '
  ria =~ ria_1 + ria_2 + ria_3 + ria_4 + ria_5 + ria_6 + ria_7 + ria_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Does not show sufficient fit.

\newpage
### EFA 1
#### Parallel analysis

```{r}
fa.parallel(d_tmp)
```

\newpage
#### Factor analysis

```{r}
efa <- fa(d_tmp, nfactors = 2, fm = "ml")
efa$loadings
```

Produces a plausible solution.

\newpage
### CFA 2

```{r}
model <- '
  ria_one =~ a*ria_2 + a*ria_4 + ria_5 + a*ria_6 + a*ria_7 + a*ria_8
  ria_two =~ ria_1 + ria_3 + ria_5
  ria_gen =~ ria_1 + ria_2 + ria_3 + ria_4 + ria_5 + ria_6 + ria_7 + ria_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Does not produce a good-fitting model. Try single-dimension next.

\newpage
### CFA 3

```{r}
model <- '
  ria_gen =~ ria_2 + ria_4 + ria_6 + ria_7 + ria_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Shows an adequate solution.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage 
## Traditionalism

```{r}
d_tmp <- d_tra
d_tmp_noninvert <- d_tra_noninvert
name_tmp <- "tra"
```

### Items

```{r}
ggplot(d_tmp_noninvert %>% 
         gather(name, value) %>% 
         left_join(items, "name"),
         aes(x = value)) +
  geom_bar() +
  facet_wrap(~content_breaks)

# extract mean and sd
desc <- data.frame(
  m = mean(unlist(d_tmp), na.rm = TRUE),
  sd = sd(unlist(d_tmp), na.rm = TRUE)
)
assign(paste0(name_tmp, "_desc"), desc)
```

\newpage
### CFA 1

```{r}
model <- '
  tra_gen =~ tra_1 + tra_2 + tra_3 + tra_4 + tra_5 + tra_6 + tra_7 + tra_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
```

Data show poor fit to the model.

\newpage
### EFA 1
#### Parallel analysis

```{r}
fa.parallel(d_tmp, fm = "ml")
```

\newpage
#### Factor analysis

```{r}
efa <- fa(d_tmp, nfactors = 3, fm = "ml")
efa$loadings
# tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Does not show a convincing solution.

\newpage
#### Factor analysis

```{r}
efa <- fa(d_tmp, nfactors = 2, fm = "ml")
efa$loadings
# tmp <- omega(d_tmp, nfactors = 2, fm = "ml")
```

Implies a single dimension, as on factor 2 there is only one significant loading.

\newpage
### CFA 2

```{r}
model <- '
  tra_gen =~ tra_1 + tra_3 + tra_5 + tra_7 + tra_8
'
fit <- cfa(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
assign(paste0(name_tmp, "_gen"), 
       dplyr::select(as.data.frame(predict(fit)), ends_with("_gen")))
```

Shows adequate fit.

```{r}
# Extract error variance
rel <- reliability(fit)["alpha", paste0(name_tmp, "_gen")]
var <- var(as.data.frame(predict(fit))[paste0(name_tmp, "_gen")])
err_var <- (1 - rel) * var
assign(paste0(name_tmp, "_err_var"), err_var)
```

\newpage
# Results
## Structural regression model

In what follows, we present the results of the structural regression model, in which we predict the three dimensions of privacy on the basis of the aforementioned personality facets. The personality facets are modelled as latent factors with single indicators, in which the indicators' error variances are specified using the results from the CFAs.

```{r}
# combine model predicted values with regular data
d_lat <- cbind(soc_gen, itg_gen, anx_gen, ria_gen, tra_gen, pri_nee_lat)
d <- cbind(d, d_lat)

# combine error variances in table
err_vars <- cbind(itg_err_var, soc_err_var, anx_err_var, tra_err_var, ria_err_var)

# define and run model
name_tmp <- "results_threefactors"
model <- '
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_9
  pri_nee_int =~ pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_7 + pri_nee_int_9
  pri_nee_ano =~ pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2
  soc =~ soc_gen
  anx =~ anx_gen
  tra =~ tra_gen
  ria =~ ria_gen
  itg =~ itg_gen

  pri_nee_gov ~ soc + itg + anx + tra + ria + male + age + inc
  pri_nee_int ~ soc + itg + anx + tra + ria + male + age + inc
  pri_nee_ano ~ soc + itg + anx + tra + ria + male + age + inc

  soc_gen ~~ .177 * soc_gen
  itg_gen ~~ .104 * itg_gen
  anx_gen ~~ .106 * anx_gen
  tra_gen ~~ .169 * tra_gen
  ria_gen ~~ .118 * ria_gen
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
```

\newpage
## Tables
### Final items measuring need for privacy

```{r}
# fig.pos="h"
# get names of included items
items_incl <- colnames(inspect(results_threefactors_fit, 'sampstat')$cov)
pri_nee_gov_items <- parameterestimates(results_threefactors_fit) %>% 
  filter(lhs == "pri_nee_gov", op == "=~") %>% 
  select(rhs) %>% 
  unlist()
pri_nee_int_items <- parameterestimates(results_threefactors_fit) %>% 
  filter(lhs == "pri_nee_int", op == "=~") %>% 
  select(rhs) %>% 
  unlist()
pri_nee_ano_items <- parameterestimates(results_threefactors_fit) %>% 
  filter(lhs == "pri_nee_ano", op == "=~") %>% 
  select(rhs) %>% 
  unlist()

t_pri_nee <- items %>% 
  filter(grepl("pri_nee", name)) %>% 
  mutate(name = factor(name, c(paste0("pri_nee_gen_", rep(4:1)),
                             paste0("pri_nee_int_", rep(9:1)),
                             paste0("pri_nee_soc_", rep(9:1))))) %>% 
  mutate(included = ifelse(name %in% items_incl, "yes", "no")) %>% 
  filter(included == "yes") %>%
  arrange(desc(as.numeric(name))) %>% 
  dplyr::select(name, content)

knitr::kable(t_pri_nee) %>% 
#  kableExtra::column_spec(2, width = "10cm") %>% 
kableExtra::kable_styling(latex_options="scale_down", font_size = 10)
```

\newpage
### Psychometrics

```{r}
t_facval <- rbind(
  "(Combined)" = fit_tab(pri_nee_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Government" = fit_tab(pri_nee_gov_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Interpersonal" = fit_tab(pri_nee_int_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Anonymity" = fit_tab(pri_nee_ano_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Sociability" = fit_tab(soc_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Integrity" = fit_tab(itg_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Anxiety" = fit_tab(anx_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Risk avoidance" = fit_tab(ria_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE),
  "Traditionality" = fit_tab(tra_fit, as_text = TRUE, reliability = TRUE, scaled = TRUE)
)

pri_nee_gov_desc <- c(mean = mean(as.matrix(d[pri_nee_gov_items])),
                      sd = sd(as.matrix(d[pri_nee_gov_items])))
pri_nee_int_desc <- c(mean = mean(as.matrix(d[pri_nee_int_items])),
                      sd = sd(as.matrix(d[pri_nee_int_items])))
pri_nee_ano_desc <- c(mean = mean(as.matrix(d[pri_nee_ano_items])),
                      sd = sd(as.matrix(d[pri_nee_ano_items])))

t_desc <- rbind(pri_nee_desc, pri_nee_gov_desc, pri_nee_int_desc, pri_nee_ano_desc,
                soc_desc, itg_desc, anx_desc, ria_desc, tra_desc)

t_psyc <- cbind(t_desc, t_facval) %>% 
  mutate_at(vars(df), funs(as.integer(.))) %>% 
  set_rownames(rownames(t_facval))
kable(t_psyc)
```

\newpage
### Structural regression model

```{r}
t_results_threefactors <- coeffs_tab(results_threefactors_fit, as_text = TRUE) %>% 
   mutate(Outcome = recode(Outcome, 
         `pri_nee_gov` = "Privacy need government",
         `pri_nee_int` = "Privacy need interpersonal",
         `pri_nee_ano` = "Privacy need anonymity")) %>% 
  mutate(Predictor = recode(Predictor,
         `soc` = "Sociability",
         `itg` = "Integrity",
         `anx` = "Anxiety",
         `ria` = "Risk avoidance",
         `tra` = "Traditionalism",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male"))
kable(t_results_threefactors)
```

\newpage
## Figures
### Bivariate relations

```{r fig.height=8, fig.width=8, out.width=".9\\textwidth", fig.align='center'}
d_lat %<>% 
  set_names("Sociability", "Integrity", "Anxiety", "Risk\navoidance", "Traditio-\nnalism", 
            "Privacy need\ngovernment", "Privacy need\ninterpersonal", "Privacy need\nanonymity")

p_bivar <- ggpairs(
  d_lat,
  upper = list(continuous = "cor"),
  lower = list(continuous = scat_plot)) +
  theme_bw()
print(p_bivar)
```

\newpage
### Structural regression model

Confidence intervals estimated using lavaan's standardized solution.

```{r}
# calculate CI for standardized effect 
d_results_threefactors <- standardizedsolution(results_threefactors_fit) %>% 
  as.data.frame() %>% 
  filter(op == "~") %>% 
  mutate(lhs = factor(lhs, c("pri_nee_ano", "pri_nee_int", "pri_nee_gov"))) %>%  # reorder for plot
  mutate(rhs = factor(rhs, c("soc", "itg", "anx", "ria", "tra", "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = recode(lhs, 
         `pri_nee_gov` = "Privacy need\ngovernment",
         `pri_nee_int` = "Privacy need\ninterpersonal",
         `pri_nee_ano` = "Privacy need\nanonymity")) %>% 
  mutate(rhs = recode(rhs,
         `soc` = "Sociability",
         `itg` = "Integrity",
         `anx` = "Anxiety",
         `tra` = "Traditio-\nnalism",
         `ria` = "Risk\navoidance",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>% 
  rename(beta = est.std)

(p_results_threefactors <- ggplot(d_results_threefactors, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) + 
  theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = c(-.4, -.2, 0, .2, .4)) +
  coord_cartesian(xlim = c(-.5, .5)) +
  facet_grid(~ rhs))
```

\newpage

Confidence intervals estimated using Bootstrap approach with 2000 draws. Note that the results differ slightly from the regular approach, which is why in the paper we present the results using lavaan's standardized solution.

```{r cache = T}
# CIs are calculated by means of bootstrapping
RNGkind("L'Ecuyer-CMRG")  # needed in order to enable replication of bootstrap
results_threefactors_boot <- bootstrapLavaan(results_threefactors_fit, 
                                             R = 2000, type = "parametric", iseed = 170819)
ci_boot <- results_threefactors_boot %>%
  as.data.frame() %>%
  select(starts_with("pri_nee_gov~"), starts_with("pri_nee_int~"), starts_with("pri_nee_ano~")) %>%
  select(-contains("~~")) %>%
  summarise_all(funs("ci_ll_boot" = quantile(., .025),
                     "ci_ul_boot" = quantile(., .975),
                     "beta_median" = median(.)))

ci_ll_boot <- select(ci_boot, contains("_ll_")) %>% t()
ci_ul_boot <- select(ci_boot, contains("_ul_")) %>% t()
beta_median <- select(ci_boot, contains("beta_median")) %>% t()

# calculate CI for standardized effect 
d_results_threefactors %<>%
  mutate(ci_ll_boot = ci_ll_boot,
         ci_ul_boot = ci_ul_boot,
         beta_median = beta_median)

(p_results_threefactors_boot <- ggplot(d_results_threefactors, aes(x = beta_median, y = lhs)) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci_ll_boot, xmax = ci_ul_boot, height = .2)) + 
  theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = c(-.4, -.2, 0, .2, .4)) +
  coord_cartesian(xlim = c(-.5, .5)) +
  facet_grid(~ rhs))
```

\newpage
# Additional analyses
## Structural regression model with need for privacy measured as bifactor
### Results

```{r}
name_tmp <- "results_bifactor"
model <- '
  pri_nee_gen =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_9 + pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_7 + pri_nee_int_9 + pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2
  pri_nee_gov =~ pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_9
  pri_nee_int =~ pri_nee_gen_1 + pri_nee_int_2 + pri_nee_int_7 + pri_nee_int_9
  pri_nee_ano =~ pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_int_2

  soc =~ soc_gen
  itg =~ itg_gen
  anx =~ anx_gen
  ria =~ ria_gen
  tra =~ tra_gen

  pri_nee_gen ~ soc + itg + anx + ria + tra + male + age + inc
  pri_nee_gov ~ soc + itg + anx + ria + tra + male + age + inc
  pri_nee_int ~ soc + itg + anx + ria + tra + male + age + inc
  pri_nee_ano ~ soc + itg + anx + ria + tra + male + age + inc

  pri_nee_gen ~~ 0*pri_nee_gov + 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_gov ~~ 0*pri_nee_int + 0*pri_nee_ano
  pri_nee_int ~~ 0*pri_nee_ano

  soc_gen ~~ .177 * soc_gen
  itg_gen ~~ .104 * itg_gen
  anx_gen ~~ .106 * anx_gen
  ria_gen ~~ .118 * ria_gen
  tra_gen ~~ .169 * tra_gen
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE, rsquare = TRUE)
assign(paste0(name_tmp, "_fit"), fit)
```

\newpage
### Visualization

```{r}
# calculate CI for standardized effect
d_results_bifactor <- standardizedsolution(results_bifactor_fit) %>%
  as.data.frame() %>%
  filter(op == "~") %>%
  mutate(lhs = factor(lhs, c("pri_nee_ano", "pri_nee_int", "pri_nee_gov", "pri_nee_gen"))) %>%  # reorder for plot
  mutate(rhs = factor(rhs, c("soc", "itg", "anx", "ria", "tra", "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = recode(lhs,
         `pri_nee_gen` = "Privacy need\ngeneral",
         `pri_nee_gov` = "Privacy need\ngovernment",
         `pri_nee_int` = "Privacy need\ninterpersonal",
         `pri_nee_ano` = "Privacy need\nanonymity")) %>%
  mutate(rhs = recode(rhs,
         `soc` = "Sociability",
         `itg` = "Integrity",
         `anx` = "Anxiety",
         `ria` = "Risk\navoidance",
         `tra` = "Traditio-\nnalism",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>%
  rename(beta = est.std)

(p_results_bifactor <- ggplot(d_results_bifactor, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) +
  theme_bw() +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = c(-.4, -.2, 0, .2, .4)) +
  coord_cartesian(xlim = c(-.5, .5)) +
  facet_grid(~ rhs))
```

\newpage
## Structural regression model with individual items

```{r}
name_tmp <- "results_items"
model <- '
  pri_nee_gen_1 + pri_nee_gen_2 + pri_nee_gen_3 + pri_nee_gen_4 + pri_nee_soc_1 + pri_nee_soc_2 + pri_nee_soc_3 + pri_nee_soc_4 + pri_nee_soc_5 + pri_nee_soc_6 + pri_nee_soc_7 + pri_nee_soc_8 + pri_nee_soc_9 + pri_nee_int_1 + pri_nee_int_2 + pri_nee_int_3 + pri_nee_int_4 + pri_nee_int_5 + pri_nee_int_6 + pri_nee_int_7 + pri_nee_int_8 + pri_nee_int_9 ~ itg_gen + soc_gen + anx_gen + tra_gen + ria_gen + male + age + inc

  itg_gen ~~ soc_gen + anx_gen + tra_gen + ria_gen + male + age + inc
  soc_gen ~~ anx_gen + tra_gen + ria_gen + male + age + inc
  anx_gen ~~ tra_gen + ria_gen + male + age + inc
  tra_gen ~~ ria_gen + male + age + inc
  ria_gen ~~ male + age + inc
  male ~~ age + inc
  age ~~ inc
'
fit <- sem(model, d, estimator = "MLM")
summary(fit, fit = TRUE, std = TRUE)
assign(paste0(name_tmp, "_fit"), sem(model, d))
```

\newpage
### Visualization

```{r}
# calculate CI for standardized effect 
d_results_items <- standardizedsolution(results_items_fit) %>% 
  as.data.frame() %>% 
  filter(op == "~") %>% 
  mutate(rhs = factor(rhs, c("soc_gen", "itg_gen", "anx_gen", "ria_gen", "tra_gen", 
                             "male", "age", "inc"))) %>%  # reorder for plot
  mutate(lhs = factor(lhs, c(paste0("pri_nee_gen_", rep(4:1)),
                             paste0("pri_nee_int_", rep(9:1)),
                             paste0("pri_nee_soc_", rep(9:1))))) %>%  # reorder for plot
  mutate(rhs = recode(rhs,
         `soc_gen` = "Sociability",
         `itg_gen` = "Integrity",
         `anx_gen` = "Anxiety",
         `ria_gen` = "Risk\navoidance",
         `tra_gen` = "Traditio-\nnalism",
         `inc` = "Income",
         `age` = "Age",
         `male` = "Male")) %>% 
  rename(beta = est.std)

(p_results_items <- ggplot(d_results_items, aes(x = beta, y = lhs)) +
  geom_vline(xintercept = 0, color = "darkgrey") +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "grey") +
  geom_vline(xintercept = -0.1, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = ci.lower, xmax = ci.upper, height = .2)) + 
  theme_bw() + 
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  scale_x_continuous(breaks = seq(-.7, .7, .2)) +
  facet_grid(~ rhs))
```

```{r save-workspace}
# remove large objects
remove(d_imp, d_incl_susp, fit)

# save workspace so that manuscript can be reproduced without rerunning analyses
save.image("../data/workspace.RData")
```